\section{Introduction}\label{sec_intro}

Modern cities struggle to build housing at the scale required to sustain growth. Rising housing costs have become one of the central economic challenges facing cities in the United States and much of the Western world.\footnote{\citet{klein2025abundance} have recently brought the issue to popular attention in their book \emph{Abundance}.} A large body of work in urban economics documents that stringent land-use regulations substantially constrain housing supply, contributing to high prices, reduced construction, and spatial misallocation of labor and capital (e.g., \citealp{glaeser2009, turner2014, hilber2016, ganongshoag2017, albouy2018, hsieh2019, brueckner2020, gabriel2020}; see \citealp{gyourkomolloy2015} and \citealp{molloy2020} for reviews). The implications extend well beyond housing markets, shaping regional inequality, migration patterns, and aggregate productivity.

Despite this growing literature, most empirical work treats regulation as a static constraint---a set of zoning rules, density limits, or statutory requirements that determine what can be built and where. Much less is known about the \textit{regulatory production process} itself---how development proposals are evaluated, modified, delayed, or approved by administrative bodies, and how this process affects urban outcomes over and above the formal text of land-use regulations. Yet in practice, housing supply is mediated through discretionary review and appeals, which require administrative processing. Even when projects ultimately comply with zoning rules or are eventually approved, conditions introduced during the administrative process can slow the pace of construction. Recent evidence suggests that prolonged approval timelines alone can significantly depress housing production (\citealp{gabrielkung2025}), highlighting the importance of bureaucratic capacity as a potential bottleneck in urban development.

This paper studies the regulatory production process in the context of a large urban planning agency. We examine decision-making by the Los Angeles City Planning Commission (LA CPC), the administrative body responsible for approving large development projects in the City of Los Angeles. Commissioners act on behalf of the city, charged with advancing development while ensuring compliance with zoning rules and public expectations. They evaluate proposals combining legal, architectural, and community considerations under time constraints. The public nature of the process generates a rich paper trail: every meeting produces agendas, minutes, and written comments that record both the information presented and the decisions made. %These records provide a window into the regulatory production process, which we seek to describe in this paper. 
We exploit this paper trail to construct a novel dataset of CPC hearing decisions. We leverage large language models to read and encode the documents, converting the over 29,000 pages of CPC documents into a dataset of \gn{NumberOfCases} decisions spanning May 2018 to February 2026, and recording the outcome of each case, the votes cast, the physical, geographic, and planning characteristics of each proposal, and the volume and content of public support and opposition.


%This paper's goal is primarily descriptive. We document the distribution of planning outcomes at the LA CPC (approval, approval with conditions, denial, and delay), as well as the factors most associated with each outcome. We examine factors including the physical characteristics of the proposed project (square footage, height, number of units, etc.), planning characteristics (requested entitlements, discretionary review), and public support or opposition. We also examine the role of bureaucratic capacity by considering factors such as the number of agenda items on the docket, the order in which the items are considered, and the atypicality of each case which reflects the cognitive burden required of commissioners to evaluate them.

%We find that project size, public opposition, and case atypicality are negatively associated with case outcome. That is---larger projects, projects receiving a larger amount of public opposition, and more atypical projects are less likely to be approved outright, and more likely to be approved with conditions, or delayed or denied. [Quantitative findings]. [Implications].

A first finding is that outright denial is extremely rare. Out of the \gn{NumberOfCases} cases in our data, only \gn{CasesDenied} were denied outright. Another \gn{CasesWithdrawn} were withdrawn, and \gn{CasesApproved} were approved outright and \gn{CasesApprovedConds} were approved with conditions or modifications. The remaining \gn{CasesContinued} were delayed to a future hearing. Out of the \gn{CasesApprovedConds} approved with conditions, \gn{CasesApprovedMajor} were classified as major conditions while \gn{CasesApprovedMinor} were classified as minor. For projects that were delayed but eventually voted on in our dataset, we observed an average of \gn{DaysToResolve} days between the first delay and the eventual resolution. Moreover, over 90\% of these projects were eventually approved. The findings suggest that delay, rather than denial or significant project alterations, is the most salient outcome hindering development at the CPC. Moreover, these delays are likely to have meaningfully large impacts on total real estate development. \citet{gabrielkung2025}, for example, estimated that if multifamily approval times had been 25\% faster in Los Angeles between 2010 and 2022, then multifamily housing production would have increased by 12.7\% in that period.

%This is not to say that regulations have no effect on the mix of projects that get built. Indeed, regulations can have a substantial effect on what kinds of projects get submitted. Our findings only indicate that \emph{once submitted}, the approvals process in Los Angeles at the CPC level does not tend to substantially alter or deny those projects but rather delays them. The regulatory approvals process and associated delays can therefore be thought of as a technical cost or friction of implementing the land-use priorities spelled out in the City's zoning code. We therefore do not claim that the regulatory approvals process is \emph{unnecessary}, just as one would not claim that criminal trials are unnecessary even if the vast majority of them result in conviction. We do claim, however, that the bureaucratic and time cost of implementing land use regulations is high and must therefore be factored into any holistic cost-benefit analysis of said regulations. 

This is not to say that regulations have no effect on the mix of projects that get built; they can substantially shape what gets submitted in the first place. Once a project lands on the CPC's desk, however, our findings show that delay is the most salient friction. We therefore turn to investigating the factors that drive approval and delay at the CPC. To do this, we develop and estimate an ordered logit model where the three outcomes are outright approval, approval with conditions, and delay or denial. We are particularly interested in teasing out the marginal influence of the project's physical characteristics, planning characteristics, public support and opposition, and bureaucratic or cognitive frictions. We emphasize the latter because the literature on bureaucratic decision-making has highlighted the importance of both principal-agent problems and bounded rationality in limiting the efficiency of the process.\footnote{On principal-agent, reputational, and incentive issues, see \citet{Dixit2002, Prendergast2003, BesleyGhatak2003, Black2008, CarpenterKrause2012, DamonteDunlopRadaelli2014, BusuiocLodge2016, MaggettiPapadopoulos2018}. For bounded rationality, limited attention, and such, see \citet{Radner1993, VanZandt1999, Sims2003, Guth2010, Salant2011, LimMatrosTurocy2014, ZhangLevin2017, deClippelRozen2021, Richters2021, HebertWoodford2023}.} Distinguishing between the influence of physical and planning characteristics from the influence of bureaucratic, political, and cognitive frictions is important because they have very different policy implications. If physical and planning characteristics are the primary drivers of delay at the CPC, then policies to speed up approvals would emphasize reforms to the zoning code and perhaps better project pre-planning. On the other hand, if bureaucratic, political, and cognitive frictions are the primary drivers of delay at the CPC, then policies to speed up approvals would emphasize public outreach, procedural streamlining, and expanding bureaucratic capacity.

After estimating the ordered logit model, we found that four factors were statistically significant and robust in predicting approval and delay. First, we found that the amount of public opposition significantly reduces the likelihood of approval and increases the likelihood of delay. A doubling in the number of opposition letters predicts a \gn{NumOpposeCoef} increase in the log odds of delay. Second, we found that the ordering of the item in the agenda significantly influences the chances of approval or delay. A project which is listed one item later in the agenda has a \gn{AgendaOrderCoef} increased log odds of approval. This is consistent with both decision fatigue and non-random ordering of agenda items (more controversial items tend to be front-loaded, as a courtesy to the public who may wish to speak on the matter), though we do not disentangle between the two. Third, we found that being placed on the consent calendar significantly predicts approval. An item on the consent calendar was found to have a \gn{ConCalCoef} increased log odds of approval. Items are placed on the consent calendar when the chair has deemed an item to be routine and non-controversial. Given that the consent calendar is a procedural tool meant to streamline decision-making, its usage and influence suggests that bureaucratic capacity is a salient constraint faced by the Commission. 

Lastly, we found that ``unusual'' cases are less likely to be approved and more likely to be delayed. To measure how unusual a case is, we develop a measure of atypicality based on how far a case is to its nearest cluster of cases in embedding space. Cases that are ``far away'' from the nearest cluster are more unusual and cases that are ``close'' to the nearest cluster are more typical. We describe this procedure in more detail in Section \ref{sec_descriptives}. We found that a one standard deviation increase in atypicality increases the log odds of delay by \gn{AtypicalityCoef}. The effect is stronger when the share of commissioners who have never heard the case before is larger, and moreover atypicality does \emph{not} significantly predict eventual denial or CPC-mandated alterations to the project, which suggests that unfamiliarity is the driver of  delay, not latent proposal quality.

%the friction it faces is delay. We see this as evidence that the bureaucratic and time costs of land-use regulation are high, and we ask which cases get slowed, and why. Part of the answer lies in the projects themselves: larger projects and projects attracting greater public opposition are less likely to be approved outright. But part of the answer lies in the body reviewing the projects: some cases are the kind the Commission has seen many times before, and some are unfamiliar. Nothing in the documents records this distinction, but it can be recovered from the relations among them. Embedding each case's description in a high-dimensional vector space, we measure how far a proposal lies from the Commission's accumulated experience, a case-level measure of atypicality. In our preferred ordered logit specification, controlling for case type, requested entitlements, embedding cluster, and council district, a one-standard-deviation increase in case atypicality reduces the log odds of full approval by \gn{AtypicalityCoef}. 
%also physical project characteristics, right?

%At its core, the LA CPC's task reflects a classic principal-agent problem shaped by two binding constraints. First, commissioners face oversight risk. Decisions are scrutinized by residents, community organizations, and the media, and public opposition can trigger political backlash. We capture this dimension by applying sentiment analysis to public correspondence as an indicator of monitoring pressure. Second, they face cognitive constraints. Development proposals vary widely in complexity and novelty, and evaluating unfamiliar cases requires greater interpretive effort. When administrative attention is scarce, routine or familiar projects may be easier to process than idiosyncratic ones. We capture this dimension by constructing a measure of case atypicality that quantifies how unusual a development proposal is relative to other cases. Using high-dimensional text embeddings of agenda items, we cluster proposals into comparable categories and measure how far each project lies from the center of its cluster. Projects with greater textual distance are more unusual and impose higher cognitive demands on commissioners. 

%The \textit{process}, as much as the projects, shapes approval --- a pattern one would expect from commissioners who act as agents on behalf of the city under two constraints. First, they face oversight risk: decisions are scrutinized by residents, community organizations, elected officials, and the media, and opposition letters are, in this sense, not merely expressions of constituent concern but signals of potential blowback. Second, they face cognitive and time constraints: dockets are long, proposals vary widely in complexity and novelty, and evaluating an unfamiliar case requires greater interpretive effort than processing a routine one. When administrative attention is scarce, familiar projects are simply easier to approve. We find this bureaucratic capacity story the most parsimonious reading of our findings, and we develop this interpretation further when discussing the results in Section \ref{sec_results}.

%We use this framework to estimate how oversight risk and cognitive constraints relate to approval outcomes using an ordered logit specification. The results reveal a systematic pattern. Less familiar projects face a penalty: controlling for project type, council district, and other case characteristics, a one-standard-deviation increase in case atypicality reduces the log odds of full approval by about \gn{SemUniCoef}. Public opposition matters as well: doubling the number of opposition letters reduces the log odds of approval by \gn{NOppCoef}, while support letters provide little offsetting benefit. In contrast, routine projects move more smoothly through the process. Placement on the Commission's consent calendar---a procedural device that bundles standard, non-controversial items for joint approval---nearly quadruples the probability of full approval. These findings provide evidence on a key mechanism through which bureaucratic performance affects housing supply. Regulatory outcomes are shaped not only by statutory rules or political incentives, but also by limits on administrative capacity to process information at scale. In this setting, bureaucratic capacity acts as an implicit constraint on development, reinforcing delay and discouraging innovation in project design. 

%The paper contributes to two strands of the literature. First, it extends research in urban economics on land-use regulation and housing supply by opening the ``black box'' of the regulatory production process. We show that the pace of development depends not only on statutory constraints but also on the administrative capacities of the regulatory apparatus itself. Second, the paper contributes to the literature on bureaucratic decision-making. At its core, the LA CPC's task reflects a classic principal-agent problem shaped by two binding constraints. First, commissioners face oversight risk. Decisions are scrutinized by residents, community organizations, elected officials, and the media. Second, they face cognitive and time constraints. Development proposals vary widely in complexity and novelty, and evaluating unfamiliar cases requires greater interpretive effort. When administrative attention is scarce, routine or familiar projects may be easier to process than idiosyncratic ones. We therefore contribute to this literature by providing field-based evidence of how bounded rationality and oversight risk shapes real regulatory outcomes in a bureaucratic setting.

The paper contributes to two strands of the literature. First, it extends research in urban economics on land-use regulation and housing supply by opening the ``black box'' of the regulatory production process. Our results suggest that the pace of development depends not only on statutory constraints but also on the administrative capacity of the regulatory apparatus itself. Second, the paper contributes to the literature on bureaucratic decision-making. Experimental studies have documented satisficing and cognitive constraints in stylized laboratory settings, but systematic evidence from the field remains scarce, in part because cognitive limits are inherently unobservable. We contribute large-scale, field-based evidence consistent with bounded rationality and oversight risk shaping real regulatory outcomes, along with a scalable, text-based approach to measuring the cognitive dimension of casework that can be applied to other institutional settings where a written record of decision-making exists.

The remainder of the paper proceeds as follows. 
%Section \ref{sec_model} presents a simple model of bureaucratic decision-making that formalizes the trade-off between cognitive effort and oversight risk. 
Section \ref{sec_background} describes the Los Angeles City Planning Commission and its role in LA's planning process. Section \ref{sec_data} describes how we collected and extracted the data used for this research. Section \ref{sec_descriptives} provides descriptive statistics regarding the data. Section \ref{sec_methodology} investigates which factors are most predictive of CPC outcomes using multivariate regression. Section \ref{sec_conclusion} concludes.


