\section{Institutional Background} \label{sec_background}

\subsection{The Los Angeles City Planning Commission}

The Los Angeles planning and approvals process for urban development is a multi-layered process that requires the input of multiple agencies. If a development is ``by-right'', meaning that it conforms to all zoning regulations within its applicable zone, then development can proceed with a permit from the Department of Building and Safety. If, however, the planned development requires an exception to the zoning code, or has a planned purpose not normally allowed by the zoning code, then approval needs to be obtained from the Planning Department and potentially from the City Planning Commission (CPC).

The LA CPC is a nine member administrative body with members appointed by the Mayor of LA and confirmed by the City Council. The CPC is responsible for reviewing citywide zoning changes, handling approvals and conditional use permits for large developments, and handling appeals of decisions made by lower bodies. The LA CPC is the highest planning body within the City's Planning Department, and sits below only the City Council in the City's overall planning hierarchy. CPC appellate decisions are often final and not appealable, and when they are appealable they may only be appealed to the City Council. 

The CPC meets approximately every two weeks. The agenda for each meeting is set by the Commission itself and made available to the public at least seven days before each meeting. To streamline hearings, the chair may place items deemed non-controversial on a consent calendar, which is voted on as a single motion. After each meeting, the meeting minutes are published online. The minutes record the order of discussion, the members present, the motions made on each agenda item, and the votes made on each motion. Motions are passed by majority rule, subject to a quorum of five members present. In addition, members of the public are allowed to submit comments on any agenda item either in writing or in person during the meeting. These procedures determine not only what the Commission decides but the conditions under which it decides: publicly, on a fixed schedule, and with its full docket in view.

\subsection{The Commission's Decision Problem} 

These conditions cast commissioners in a decision environment where they act as agents on behalf of the city. They are appointed rather than elected, they serve part-time, and they are charged with advancing development while ensuring compliance with zoning rules and public expectations. The capacity and conduct of individual bureaucrats in positions like this have measurable consequences for government performance \citep{decarolisetal2020, fenizia2022, besthjortszakonyi2023}, and overloading bureaucracies relative to their capacity degrades the quality of implementation \citep{grattonetal2021}. The cases that reach the CPC are by construction those that the zoning code does not resolve on its own, so the discretion commissioners exercise is genuine. But they exercise it in public, on the record, and subject to two constraints.

The first constraint is oversight risk. Every agenda, motion, and vote is published. Hearings are attended by residents, community organizations, and representatives of applicants, and CPC decisions can be appealed to or vetoed by the City Council. Oversight of this kind operates through what \citet{mccubbinsschwartz1984} call fire alarms, in which affected parties raise the alert when a decision threatens their interests. Written public comments are the fire alarms of the planning process, and the people who pull them are disproportionately opposed to new development and effective at delaying them  \citep{einsteinglickpalmer2019, hankinson2018}. The resulting exposure is asymmetric. A commissioner who approves a contested project makes an attributable decision, whereas postponing the same project makes no such commitment. None of this requires commissioners to be careless or self-interested. Caution of this kind can be an equilibrium response to the incentives bureaucrats face \citep{prendergast2003}.

The second constraint is attention. The Commission meets roughly biweekly and typically faces around ten decision items per hearing, which vary enormously in the interpretive effort they demand. Many have been seen by the Commission before, and can be evaluated largely on pattern. Others present questions the Commission has no template for, such as a request to host concerts, film screenings, and charitable fundraisers, with alcohol service, on the grounds of a 53-acre cemetery (a real case in our data, to which we return in Section~\ref{sec_descriptives}). An unfamiliar case must be worked through from scratch, and the time spent doing so is time not available for the rest of the docket. The consequences of scarce attention in screening bureaucracies are well documented. Patent examiners given less time per application scrutinize less and grant more invalid patents \citep{frakeswasserman2017}, and professional decision makers ranging from asylum judges to loan officers exhibit systematic cognitive errors when processing long sequences of cases \citep{chenmoskowitzshue2016}. Decision makers are boundedly rational \citep{simon1955}, allocating scarce attention across competing demands \citep{mackowiaketal2023}. What remains scarce is evidence on how these constraints operate inside administrative bodies, where the cognitive demands of a case are not directly observable.

A body operating under these two constraints would be expected to economize on both attention and exposure. Familiar, uncontested cases should move quickly. The Commission has in fact built a device for this purpose, the consent calendar, which batches items the chair deems non-controversial into a single vote. How the Commission
handles the rest of its docket---which cases move quickly, which acquire
conditions, and which are put off---is not something its rules announce; it
must be read off the record of its decisions. The two constraints described
here supply the lens through which we read that record, and we return to
them when interpreting the results in Section~\ref{sec_methodology}.