\section{Data Description} \label{sec_descriptives}

The resulting dataset contains \gn{NumberOfCases} agenda items with motions voted on by the CPC. Because the dataset is novel, we devote an entire section to describing and exploring the data. These descriptive results may be useful for future researchers who want to work with similar data, or for policymakers who want to better understand the current planning process in Los Angeles.

\subsection{Distribution of project outcomes}

Table \ref{tab_result_unanimity} shows the distribution of the project outcomes by the unanimity of the vote. Two important facts emerge about the CPC process. First, a minuscule number of projects are denied outright (\gn{CasesDeniedPct}\% of all cases). Rather than being denied, a more common occurrence is that the decision is postponed to a later date (\gn{CasesContinuedPct}\% of cases) or the requested actions are approved with major or minor conditions or modifications  (\gn{CasesApprovedMajorPct}\% and \gn{CasesApprovedMinorPct}\% repsectively).\footnote{Examples of minor conditions and modifications include minor adjustments to the approved Floor Area Ratio, minor changes in number of approved parking spaces, and adding required security camera locations. Examples of major conditions and modifications include major changes to the allowed number of dwelling units, increases to the required number of affordable set-asides, and changes to the requested zoning designation.} \gn{CasesApprovedPct}\% of the cases were approved without any further conditions or modifications. 

Second, most decisions were unanimous (\gn{CasesUnanimousPct}\% of all cases). Because of these two facts, we do not view disagreement \emph{within} the board as a significant source of friction in moving development projects through the pipeline. Moreover, few projects are denied outright, so the impact of CPC hearings on final outcomes must come through either i) a slowing down of the process due to having to wait for the decision, which itself may be delayed multiple times; or ii) CPC-mandated alterations that potentially add cost or time to the development, or which could  dissuade the developer from even moving forward with the project. Our primary analysis will therefore focus on the bureaucratic factors which lead to either delayed decision making or attaching conditions to the project proposal.

\subsection{Distribution of public comments}

Of the \gn{NumberOfSupplementalDocs} supplemental documents included in the dataset, \gn{NumberOfPublicComments} were identified as public comments. \gn{NumberInSupport} expressed support for one or more project proposals on the agenda to which they were submitted, and \gn{NumberInOpposition} expressed opposition to one or more project proposals.\footnote{The numbers expressing support and opposition do not necessarily add up to the total number of public comments because some letters may be neutral, while others may express support for some agenda items and opposition to others.}  The average number of public comments received for the \gn{NumberOfCases} cases in our analytical dataset was \gn{AvgLettersPerCase} letters per case.\footnote{The number of letters per case is not simply the number of public comments divided by the number of cases because some letters address multiple projects. We count such letters as separate instances of a public comment for each project the letter references.} Figure \ref{fig_support_oppose} shows the distribution of the number of letters received per case, in support and in opposition. The distribution is thick-tailed with he majority of cases receiving under 5 letters, whereas a smaller number of controversial cases receive the lion's share of public comments.

Table \ref{tab_reasons} documents the reasons offered for support or opposition in public comments. For each letter supporting or opposing an agenda item, the LLM was asked to extract a list of reasons for the support or opposition, chosen from a list of possible categories. Table \ref{tab_reasons} shows the possible categories, and the number of letters mentioning each of those categories, grouped separately by support letters and opposition letters. Within each group (support or opposition), the categories are listed in order of frequency of mention, thus showing which categories are most salient to the public. 

Among support letters, the most frequently mentioned reasons are related to housing supply and affordability, transit access and walkability, and procedural issues. We note that support letters are less likely to be written by individual members of the public, and more likely to be written by representative of developers or by housing advocacy organizations, and thus the reasons for support reflect the priorities of this group of actors. The category of ``procedural issue'' relates to questions on whether the developer followed proper procedures in seeking project approval. Support letters mentioning procedural issues are offered in defense of the developer having followed proper procedures.

Among opposition letters, the most frequently mentioned reasons are related to neighborhood character, building dimensions, and procedural issues.  Opposition letters are more likely to be written by individual members of the public, and sometimes by neighborhood councils, so the reasons for opposition reflect the priority of this group of actors.  When opposition letters mention procedural issues or community engagement, the complaint is usually that the developer did not follow proper procedure or did not engage sufficiently with community concerns.

\subsection{Distribution of case characteristics}

Table \ref{tab_bivariate_descriptives} shows summary statistics for selected case characteristics. For continuous variables, the sample mean is shown both for the overall dataset and by case outcome (approved, approved with conditions, or delayed or denied). For boolean variables, the fraction of cases having that feature is shown, both for the overall sample and separately by case outcome. For each variable, a statistical test is also performed to test equality of means by case outcome and the p-value of that test is shown for each variable.\footnote{For continuous variables, a one-way ANOVA test is performed. For boolean variables, a Pearson's chi-squared test is performed.}

\paragraph{Case Type.} \gn{PctCasesResidential}\% of the cases were proposals for new residential developments, \gn{PctCasesMixedUse}\% were for new mixed-use developments (meaning some degree of residential), and \gn{PctCasesNonResidential}\% were for new non-residential developments. The remaining \gn{PctCasesOther}\% include cases such as requests for conditional use permits for existing developments, or proposals to amend the zoning code in various ways, including both Specific Plan and General Plan updates.

Residential cases are more likely to be approved. Although residential cases account for \gn{PctCasesResidential}\% of all cases, they account for \gn{PctApprovedCasesResidential}\% of all approved cases, and only \gn{PctDelayedCasesResidential}\% of delayed/denied cases. Non-residential cases are less likely to be approved. Although non-residential cases account for \gn{PctCasesNonResidential}\% of all cases, they account for \gn{PctApprovedCasesNonResidential}\% of approved cases and \gn{PctDelayedCasesNonResidential}\% of delayed/denied cases. There was no statistical evidence that mixed-use or other type cases were systematically more or less likely to be approved.

\paragraph{Physical Characteristics.} The average proposed development in our data was \gn{AvgSquareFootage} square feet large and \gn{AvgHeight} feet tall, and contained \gn{AvgDwellingUnits} dwelling units, of which \gn{AvgAffordableUnits} were deed-restricted affordable units, and \gn{AvgParkingSpaces} parking spaces. Unsurprisingly, larger projects were less likely to be approved outright, and more likely to be approved with conditions or delayed or denied. 

\paragraph{Planning Characteristics.} We identified the planning characteristics of each project by utilizing the suffixes in the Planning Department case number. Each suffix indicates a feature of the case relevant the Planning Department. For example, a suffix of SPR indicates that the development requires a Site Plan Review, and a suffix of DB indicates that the development is utilizing Density Bonus incentives. Altogether, there are \gn{NumUniqueSuffixes} unique suffixes in the dataset, which we grouped into \gn{NumSfxGroups} categories: Site Plan Review, Appeals, Incentive Programs, Conditional Use Permits, Variances/Adjustments/Exceptions, Compliance Review, Code or Plan Amendments, Development Agreements, and Other.\footnote{Refer to Appendix Tables \ref{tab_suffixes_all} and \ref{tab_suffixes_all_2} for a full list of all \gn{NumUniqueSuffixes} suffixes, their meanings, and their assigned categories.}

\emph{Site Plan Review.} This category indicates that the project needed to undergo Site Plan Review. \gn{PctSPR}\% of projects had a suffix in this category. Projects requiring Site Plan Review were less likely to be approved outright, and more likely to be approved with conditions.

\emph{Appeals.} This category indicates that the CPC is hearing an appeal related to a former decision on this project. \gn{PctAPP}\% of projects had a suffix in this category. Appealed projects are less likely to have further conditions placed on them; they are more likely to be either postponed or approved as-is.

\emph{Incentive Programs.} This category indicates that the project is utilizing or requesting to utilize an incentive program offered by the City or State. These include programs like Los Angeles's Transit Oriented Communities program, which allows for streamlined development near public transit, or California's Density Bonus law, which allows for higher density zoning in exchange for affordable unit set-asides. \gn{PctIP}\% of projects had a suffix in this category. Projects utilizing an incentive program were more likely to be approved outright and less likely to be approved with conditions or delayed.

\emph{Conditional Use Permits.} This category indicates that the development is requesting a conditional use permit. These are permits to use the property in ways that the zoning code does not normally allow for, like operating a school of a certain size, or selling alcohol.  \gn{PctCUP}\% of projects had a suffix in this category. Projects requesting a conditional use permit were less likely to be approved outright, and more likely to be approved with conditions or postponed.

\emph{Variances/Adjustments/Exceptions.} This category indicates that the development is requesting a zone variance, line adjustment or other exception to the zoning code. \gn{PctVAE}\% of projects had a suffix in this category. Projects requesting a variance, adjustment, or exception were less likely to be approved outright and more likely to be approved with conditions or delayed.

\emph{Compliance Review.} This category indicates that the development had to undergo a review for compliance with rules related to an overlay zone or other special district that applies to the property. \gn{PctCR}\% of projects had a suffix in this category. Projects requiring compliance review were more likely to be approved with conditions, and less likely to be approved outright or delayed, though the differences are not statistically significant.

\emph{Code or Plan Amendments.} This category indicates that the proposal seeks to make changes to the zoning code or the City's General Plan or an applicable Specific Plan. \gn{PctCPA}\% of projects had a suffix in this category. Projects requesting a code or plan amendment were more likely to be approved with conditions, and less likely to be delayed.

\emph{Development Agreements.} This category indicates that the project involves an agreement between the developer and the City, such as the City granting a waiver of dedication and improvements, or the City allowing one developer to transfer development rights to another developer. \gn{PctDA}\% of projects had a suffix in this category. Projects with development agreements were less likely to be approved outright and more likely to be delayed, though the differences are not statistically significant.


\paragraph{Public Input.} The average number of public comments per case was \gn{AvgNumLetters}, of which \gn{AvgNumSupport} were in support and \gn{AvgNumOppose} were in opposition. Receiving more letters is associated with greater controversy and a higher likelihood of delay. The average number of letters received by delayed/denied cases was \gn{AvgNumLettersDelayed}, whereas the average number of letters received by approved cases was \gn{AvgNumLettersApproved}. The difference is statistically significant at the 1\% level. It is interesting to note that delayed cases received both more letters in opposition \emph{and} more letters in support. This suggests that letters of support are often written only in response to the controversy that a case generates.

\paragraph{Hearing Characteristics.} We also explored whether logistical features related to the hearing itself mattered for case outcomes. First, we looked at the order of the item in the agenda. The average agenda item requiring a decision from the board appeared as the \gn{AvgAgendaOrder}th item on the agenda. In these univariate comparisons, there does not appear to be any evidence that the agenda item order matters for case outcome (though this conclusion changes as we move to the multivariate regressions). 

Second, we looked at how many items were on the agenda for that hearing. In theory, a busier agenda could mean that cases on that agenda are more likely to get postponed and less likely to be approved with conditions, since adding conditions takes extra time. We do find some evidence of this: cases on fuller agendas are more likely to be postponed and less likely to be approved with conditions, and the difference is statistically significant at the 5\% level. 

Third, we looked at the influence of an item being on the ``consent calendar''. The consent calendar is a practical tool for streamlining meetings in which multiple items are voted on as a whole in a single motion. The consent calendar includes items that the Commission chair has deemed non-controversial and therefore not requiring separate discussion. The consent calendar is published before each hearing, and items can be taken off by Commission members before or after the meeting. \gn{PctConsentCalendar}\% of the cases were voted on as part of a consent calendar, and cases on the consent calendar were significantly more likely to be approved and significantly less likely to be delayed or denied.

\paragraph{Case Atypicality.} In addition to logistical features of the hearing, we wanted to test whether the unusualness of a case matters for its outcomes. In theory, highly unusual cases demand more cognitive bandwidth, and possibly more discussion time, from Commission members. Thus, we hypothesize that decisions on highly atypical cases are more likely to be postponed than decisions on run-of-the-mill cases. 

To compute a measure of case atypicality, we first calculated the text embeddings of each agenda item using OpenAI's \texttt{text-embedding-3-small} embeddings model. A text embedding is a $N$-dimensional vector representation of a text's meaning.\footnote{For an introduction to the concept of embeddings, see \citet{mikolov2013} and \citet{le2014}.} To remove potential differences in formatting, boilerplate, or how individual CPC staff members write agenda items, we first asked \texttt{claude-opus-4-6} to provide a short summary of the item, focusing on key development aspects of the case, such as project type, location, and requested actions. This summary was then passed into the embedding model. For each agenda item, the model returned a 1,536 dimensional numerical vector that represents the semantic meaning of the text. Two agenda items with very similar proposals will have embeddings that are close to each other, while two agenda items with very different proposals will have embeddings that are far away from each other. 

Because \texttt{text-embedding-3-small} was trained on a general corpus of documents, not specialized to the topic of municipal planning and zoning, not all 1,536 dimensions may be relevant for capturing the important differences between our agenda items. We therefore used principal components analysis to extract the 10 linear combinations of these dimensions which explain the most variance in our corpus. Figure \ref{fig_scree_plot} shows the scree plot of the principal components analysis. By the 10th principal component, there are diminishing returns to including more components.

After reducing the embeddings down to 10 dimensions, we grouped the agenda items into three clusters using K-means clustering. Identifying clusters in our data is important because the CPC handles many different types of cases, and the language used could be different across different case types. For example, the language used in a case involving the demolition of an existing building followed by the reconstruction of a multifamily home would be different from a request for a conditional use permit to operate a school. Thus, we measure atypicality only within a set of similar case types, as defined by the clustering algorithm.

Figure \ref{fig_clusters} shows a scatter plot of the agenda items according to their first two principal components, colored by cluster. 
\input{results/cluster_summaries.tex}

To measure the atypicality of an agenda item, we calculated the Mahalanobis distance between the agenda item's 10-dimensional PCA-reduced embedding to its cluster's centroid. We chose Mahalanobis distance because it accounts for the correlation structure of the reduced embedding space \citep[p. 93]{rencher2012}. A simple Euclidean distance assumes that each principal component is uncorrelated and equally informative. By contrast, the Mahalanobis distance rescales and rotates the coordinate system so that correlations between dimensions are removed. This provides a scale-free and cluster-comparable measure of uniqueness. A Mahalanobis distance of 1.0 indicates that the agenda item lies one standard deviation away from its centroid, regardless of which cluster it belongs to or how dispersed that cluster is in the original embedding space.

%When computing Mahalanobis distance, we assessed the stability of the covariance matrices. For each cluster, we compared the number of agenda items to the dimensionality of the reduced embedding, finding that all clusters exceeded the conventional guideline. We also evaluated condition numbers and minimum eigenvalues, which confirmed that the covariance matrices were well-conditioned and positive definite \citep[p.~268]{aguinis2009}. 

%In addition to including the Mahalanobis distance to the cluster centroid as a measure of case atypicality, we include fixed effects for each cluster. This allows each cluster to have a separate baseline probability distribution over outcomes.

Using our measure of case atypicality, we can get a sense of which cases are most unusual for each cluster. 
\input{results/unique_examples.tex}

%\paragraph{Perplexity.} Semantic uniqueness measures how unusual a proposal is relative to other cases of similar type. In addition to that, we also try to measure how confusing or hard to understand the proposal is in general. To do this, we ask \texttt{gpt-4o} to summarize each agenda item, then we measure the perplexity of the response. In language models, perplexity is a measure of how uncertain the model is about its response. It is measured as the exponent of the response's cross-entropy  (i.e. the negative mean of the output tokens' conditional log probabilities.\footnote{See \citet{jm2}.}) A perplexity of 1 indicates that the model has no uncertainty about its output, while a higher perplexity means the model is more uncertain.  % NO LONGER USING PERPLEXITY

Table \ref{tab_bivariate_descriptives} shows how the average value of case atypicality differs across case outcomes. The averages themselves are not meaningful, but the differences across outcome are.\footnote{The average of the squared-Mahalanobis distance is always equal to the dimensionality of the feature space.} Table \ref{tab_bivariate_descriptives} shows a difference of \gn{AtypicalityDifference} between the average atypicality of a delayed/denied case and the average atypicality of an approved case, which means that delayed/denied cases are about 0.22 standard deviations more atypical than approved cases. The difference is statistically significant at the 1\% level.

\subsection{Summary of findings}

To summarize our findings in this section, we found that: i) residential developments were more likely to be approved and less likely to be delayed, whereas non-residential developments experienced the opposite; ii) larger projects were more likely to be delayed or approved with conditions, and less likely to be approved outright; iii) planning characteristics influence case outcomes, but in various non-monotonic ways; iv) cases receiving a higher amount of public scrutiny were more likely to be delayed and less likely to be approved; v) cases were more likely to be delayed and less likely to be approved with conditions when the agenda was fuller; and vi) atypical cases were more likely to be delayed and less likely to be approved outright.

The results are consistent with a model of bureaucratic decision making in which time constraints, cognitive constraints, and public monitoring are all relevant factors in determining the outcome. However, the descriptive results presented in this section do not disentangle the partial effects of each feature from the effects of other correlated features. Therefore, we develop and present the results from a multivariate model in the next section.





