\section{Statistical Model of CPC Decision-Making}\label{sec_methodology}

\subsection{Model}

To investigate the partial effects of case characteristics on case outcome, we developed and estimated a statistical model of CPC decision-making. We model each agenda item at a CPC hearing as having three possible outcomes:
\begin{enumerate}[start=0]
\item The project is denied or the decision is postponed;
\item The project is approved with conditions;
\item The project is approved without further conditions.
\end{enumerate}
Each project proposal $i$ is assumed to have a latent quality variable $y_i^\ast$ which determines the likelihood of the three outcomes. $y_i^\ast$ is modeled as a linear function of observed project characteristics and bureaucratic factors, $\mathbf{X}_i$, plus an error term $\epsilon_i$:
\begin{align}
y_i^\ast = \mathbf{X}_i \beta + \epsilon_i
\end{align}
The latent quality of the project proposal determines its outcome at the CPC hearing. Let $y_{i} \in \{0, 1, 2\}$ denote project $i$'s outcome. The relationship between $y_i^\ast$ and $y_i$ is as follows:
\begin{align}
y_i = \begin{cases}
0 \text{ if } y_i^\ast < \mu_0 \\
1 \text{ if } \mu_0 \leq y_i^\ast < \mu_1 \\
2 \text{ if } \mu_1 \leq y_i^\ast
\end{cases}
\end{align}
with $\mu_0 < \mu_1$. $\epsilon_{i}$ is assumed the follow a standard logistic distribution, and therefore the model is an ordered logit model, and the outcomes are monotonic in $\mathbf{X}_i \beta$. The parameters $\beta$, $\mu_0$, and $\mu_1$ are estimated by maximum likelihood.

\subsection{Estimation Results}

To operationalize the model, we include in $\mathbf{X}_i$ the following variables: indicators for whether the project is a residential, mixed-use, or non-residential development, the natural log of the square footage of the project, the height of the project in feet, the log-base-2 of the number of support letters, the log-base-2 of the number of opposition letters, the order in which the item appears on the agenda, the number of agenda items for that date's hearing, and the case atypicality.\footnote{We chose not to include other physical characteristics such as number of dwelling units, number of affordable units, and parking spaces because they are highly correlated with the other physical dimensions. We also include dummy variables for whether the case is missing height or square footage information.} In addition, we include an increasingly comprehensive set of dummy variables, including dummies for the case suffix categories, the Council District to which the project belongs, the year in which the case was heard, and the semantic embeddings cluster the case belongs to.

Table \ref{tab_ologit_results} reports the estimated model coefficients for four different specifications. In column 1, the model includes only dummies for case suffix category, but not for Council District, year, or embedding cluster. In column 2, the model adds Council District dummies. In column 3, the model further adds year dummies. In column 4, the model adds embedding cluster dummies. The purpose of the successive inclusion of dummies is to show robustness of the coefficients of interest to the inclusion of further controls. Column 4 is the preferred specification.

Table \ref{tab_ologit_results} shows that although physical dimensions were associated with outcomes in the univariate analysis, the association disappears when other factors are included in the regression. The coefficients on height and square footage are neither large nor statistically significant. Similarly, in the univariate analysis there was a statistically significant association between development type and outcomes. But in the multi-variable regression analysis, the association between project type and outcome is no longer statistically significant.

The variables which are revealed to have a statistically significant association with outcomes are: the number of opposition letters, the order of the item in the agenda, whether the case is on the consent calendar, and the case atypicality. Using column 4 as the preferred specification, the results show that: i) a doubling in the number of opposition letters is associated with a \gn{NumOpposeCoef} reduction in log odds of full approval; ii) moving an agenda item one spot down in the agenda is associated with a \gn{AgendaOrderCoef} increase in log odds of full approval; iii) being on the consent calendar is associated with a \gn{ConCalCoef} increase in log odds of full approval; and a one standard deviation increase in atypicality is associated with a \gn{AtypicalityCoef} decrease in log odds of full approval.

To translate these results into more interpretable numbers, we report the marginal effects of each variable in Table \ref{tab_ologit_marginal_effects}. The results indicate that for an ``average'' case: i) doubling the number of opposition letters is associated with a \gn{NumOpposeME2}\% reduction in the probability of full approval, but a \gn{NumOpposeME1}\% increase in the probability of approval with conditions, and a \gn{NumOpposeME0}\% increase in the probability of delay or denial; ii) moving an item down one spot in the agenda is associated with a \gn{AgendaOrderME2}\% increase in the probability of full approval, but a \gn{AgendaOrderME1}\% decrease in the probability of approval with conditions, and a \gn{AgendaOrderME0}\% decrease in the probability of delay or denial; iii) being on the consent calendar is associated with a \gn{ConCalME2}\% increase in the probability of full approval, and a \gn{ConCalME1}\% decrease in the probability of approval with conditions, and a \gn{ConCalME0}\% decrease in the probability of delay or denial; iv) a one standard deviation increase in atypicality is associated with a \gn{AtypicalityME2}\% decrease in the probability of full approval, but a \gn{AtypicalityME1}\% increase in the probability of being approved with conditions, and a \gn{AtypicalityME0}\% increase in the probability of being delayed or denied.

Overall, these results suggest that bureaucratic and political factors drive marginal decision-making at the CPC. If the CPC was primarily motivated by approving small residential developments and denying large non-residential ones, then we would expect to see statistically significant coefficients on project type and physical dimensions. Similarly, if the CPC was primarily motivated by strict adherence to the letter of the law in the zoning code, then we would not expect to find statistically significant influence of public sentiment and logistical issues, once planning characteristics are controlled for. Yet, that is what we find.

The signs of the coefficients are consistent with a theory of bureaucratic decision-making under time constraints, cognitive constraints, and oversight risk. Decisions on unfamiliar projects may be postponed because they require greater cognitive bandwidth, which is scarce under time constraints. Contested projects are also more likely to be postponed because they expose Commission members to greater political and reputational scrutiny. By contrast, routine and uncontested projects proceed with little resistance, as reflected by the consent calendar. 

\subsection{Bureaucratic capacity or latent quality?}

We have interpreted the effect of agenda order and atypicality as arising primarily from time constraints and cognitive constraints. That is, we have interpreted their effects as matters of bureaucratic capacity. A natural question that arises is whether their effects could instead be due to lower underlying quality.  If atypical proposals are of lower underlying quality, or if items earlier in the agenda are of lower underlying quality, then we can no longer interpret their association with delay as a matter of bureaucratic capacity, but rather as a matter of underyling quality issues with the proposal.

To test this, we ask whether delayed cases are more likely to be eventually denied. If atypical and agenda order were merely proxies for latent quality, we would expect them to predict worse eventual outcomes regardless of whether the case was initially delayed; that is, they should be significantly more likely to be denied or approved with conditions in subsequent hearings. To test this hypothesis, we remove all delay decisions from the estimation sample and estimate a binary logit model in which the outcomes are 1 if the case is approved and 0 if the case is either approved with conditions or denied. The results for this regression are shown in Table \ref{tab_blogit_results}. In all four specifications, the coefficients on atypicality and agenda order becomes substantially attenuated. The coefficients on atypicality are no longer statistically significant in any of the four specifications, whereas the coefficients on agenda order are only statistically significant in the fourth regression, and only at the 10\% level. Thus, atypicality and agenda order have more statistical power in explaining delay than they do in explaining eventual outcome. This suggests that the association of agenda order and atypicality to delay cannot be attributed entirely to poor latent quality.

\subsection{What happens to delayed cases?}

We now assess the extent to which delays at the CPC have a meaningful impact on economic outcomes. \citet{gabrielkung2025} showed that approval delays are an important driver of reduced housing production. They estimated that a 25\% reduction in approval times would have increased housing production in Los Angeles by 12.7\% between the years of 2010 and 2022. Thus, delays at the CPC could have a meaningful impact on real estate development if they add significant time to the overall approvals process. 

Table \ref{tab_delayed_outcomes} investigates what happens to the delayed cases in our dataset. There were \gn{NumDelayedCases} cases which were ever delayed in our dataset.\footnote{This number differs from the \gn{CasesContinued} instances of delay reported in Table \ref{tab_result_unanimity} because in Table \ref{tab_result_unanimity} counts multiple hearings of the same case as separate units.} Of these, \gn{NumResolved} eventually received a non-delay outcome within our dataset. The average time between first observed delay and final outcome was \gn{DaysToResolve} days. This is an underestimate because a) we don't observe the final outcome for some of the cases (right-censored resolution times) and b) we don't necessarily observe the first hearing (left-censored start times). But even this underestimate of \gn{DaysToResolve} days is economically significant: \citet{gabrielkung2025} report that the median approval time for multifamily housing development projects in Los Angeles from 2010 to 2022 was about 524 days, so \gn{DaysToResolve} days is a meaningful addition. Moreover, it is interesting to note that between the initial delay and final outcome, the majority of the delayed cases (\gn{ResolvedNoChangesPct}\%) did not change in terms of their requests, and some (\gn{ResolvedMinorChangesPct}\%) came back with only minor changes. Only \gn{ResolvedMajorChanges} of the delayed cases came back with any major changes to their requests; and even among these, only \gn{ResolvedMajorReduce} came back with reduced requests, while the other \gn{ResolvedMajorExpand} actually expanded their requests. The fact that a delay at the CPC does not usually cause the applicant to modify its requests again suggests that the CPC's decision to delay is not necessarily correlated with the quality of the proposal, or else one would expect the developers to modify the proposal to increase the chance of approval.

\subsection{Discussion}

%We have found evidence that bureaucratic and political factors are statistically significant drivers of decision-making at the CPC. We do not claim, however, that those factors matter more than the underlying quality of the project or its adherence to proper procedures and standards. It may simply be that by the time a project lands on the CPC's desk, it has already been pre-screened for quality, such that basic quality issues are no longer what determines approval at the margin. Unfortunately, we are not able to test this hypothesis with the data that we have, and we leave such work to future research. Our claim is much more modest: that decision-making at the CPC is slowed down by bureaucratic and political factors that may potentially be orthogonal to the overall planning objectives of the City. Thus, bureaucratic and political delay should register in the cost side of any cost-benefit analysis of new regulation. This is, of course, known quite generally, but we provide novel empirical evidence and some quantification of the effect.

Cognitive limits are inherently unobservable. They leave no entry in an agenda or a minute, and can only be inferred indirectly from patterns in observed decisions. Our estimates provide the basis for one such inference. If approval at the CPC hinged primarily on the fundamentals of a proposal, the fundamentals should carry the explanatory weight. Instead, project type and physical dimensions lose their association with outcomes once controls are included, while the variables describing the process---public opposition, case atypicality, and placement on the consent calendar---retain theirs.

The decision process described in Section~\ref{sec_background} offers an interpretation of this pattern, where the opposition penalty reflects oversight risk and the atypicality penalty comes from scarce attention. But an alternative reading is that these variables simply proxy for latent quality. Atypical cases may be atypical because something is wrong with them. Two features of the data resist this interpretation. First, once delayed cases are set aside, atypicality loses its power to predict final outcomes. Atypicality predicts whether a case is delayed, not how it is ultimately decided, supporting a processing friction rather than a quality defect. Second, delayed cases overwhelmingly return unchanged: roughly three quarters come back with identical requests and are then resolved. If delay signaled a quality problem, applicants would presumably revise. Instead, for the most part, they simply wait.

Nonetheless, it may be that by the time a case reaches the CPC it has been screened for basic quality, so that the frictions we document operate only at the margin among pre-vetted proposals. We cannot test these possibilities with the data at hand, and we flag them as open questions rather than resolve them. Our claim is modest: decision-making at the CPC is slowed by bureaucratic and political factors that may be orthogonal to the overall planning objectives of the City. Bureaucratic and political delay should accordingly register on the cost side of any cost--benefit analysis of new regulation. This is, of course, known quite generally, but we provide novel empirical evidence and some quantification of the effect.
















%\subsection{Explanatory Variables}

%We now turn to discussing the explanatory variables we include in the model.

%\paragraph{Case atypicality.} In that one of our goals is to quantify the role of bureaucratic frictions in the CPC approvals process, we here develop a concept which we call ``case atypicality''. At a high level, case atypicality is designed to capture how unique an agenda item is relative to other agenda items that the CPC is accustomed to hearing. Our hypothesis is that agenda items which are highly unusual may be less likely to be approved outright, more likely to be approved with conditions, and more likely to be denied or delayed. We thus expect to estimate a negative coefficient on case atypicality.

%To compute a measure of case atypicality, we first calculate the text embeddings of each agenda item using OpenAI's \texttt{text-embedding-3-small} embeddings model.\footnote{For an introduction to the concept of embeddings, see \citet{mikolov2013} and \citet{le2014}.} To remove differences in formatting, boilerplate, or how individual CPC staff members write up agenda items, we first ask \texttt{claude-opus-4-6} to provide a short summary of the item, focusing on key development aspects of the case, such as project type, location, and requested actions. This summary is then passed into the embedding model. For each agenda item, the model returns a 1,536 dimensional numerical vector that represents the semantic meaning of the text. Two agenda items with very similar proposals will have embeddings that are close to each other, while two agenda items with very different proposals will have embeddings that are far away from each other. 

%Because \texttt{text-embedding-3-small} was trained on a general corpus of documents, not specialized to the topic of municipal planning and zoning, not all 1,536 dimensions may be relevant for capturing the important differences between our agenda items. We therefore use principal components analysis to extract the 10 linear combinations of these dimensions which explain the most variance in our corpus. Figure \ref{fig_scree_plot} shows the scree plot of the principal components analysis. By the 10th principal component, there are diminishing returns to including more components.

%After reducing the embeddings down to 10 dimensions, we group the agenda items into three clusters using K-Means clustering. Identifying clusters in our data is important because the CPC handles many different types of cases, and the language used could be very different across different case types. For example, the language used in a case involving the demolition of an existing building followed by the reconstruction of a multifamily home would be very different from a request for a conditional use permit to operate a school. Thus, it only makes sense to measure atypicality within a set of similar case types.

%Figure \ref{fig_clusters} shows a scatter plot of the agenda items according to their first two principal components, colored by cluster. 
%\input{results/cluster_summaries.tex}

%To measure the atypicality of an agenda item, we calculate the Mahalanobis distance between the agenda item's 10-dimensional PCA-reduced embedding to its cluster's centroid. We chose Mahalanobis distance because it accounts for the correlation structure of the reduced embedding space \citep[p. 93]{rencher2012}. A simple Euclidean distance assumes that each principal component is uncorrelated and equally informative, and even if rescaled by per-dimension variance, it would still treat dimensions as independent. In contrast, the Mahalanobis distance rescales and rotates the coordinate system so that correlations between dimensions are removed. The estimated coefficient are interpreted as the change in log odds associated with moving one standard deviation farther from the cluster centroid. This provides a scale-free and cluster-comparable measure of uniqueness. A value of 1.0 indicates that an agenda item lies one standard deviation away from its centroid, regardless of which cluster it belongs to or how dispersed that cluster is in the original embedding space.
%When computing Mahalanobis distance, we assessed the stability of the covariance matrices. For each cluster, we compared the number of agenda items to the dimensionality of the reduced embedding, finding that all clusters exceeded the conventional guideline. We also evaluated condition numbers and minimum eigenvalues, which confirmed that the covariance matrices were well-conditioned and positive definite \citep[p.~268]{aguinis2009}. 
%In addition to including the Mahalanobis distance to the cluster centroid as a measure of case atypicality, we include fixed effects for each cluster. This allows each cluster to have a separate baseline probability distribution over outcomes.

%Using our measure of case atypicality, we can get a sense of which cases are most unusual for each cluster. 
%\input{results/unique_examples.tex}

%\paragraph{Perplexity.} Semantic uniqueness measures how unusual a proposal is relative to other cases of similar type. In addition to that, we also try to measure how confusing or hard to understand the proposal is in general. To do this, we ask \texttt{gpt-4o} to summarize each agenda item, then we measure the perplexity of the response. In language models, perplexity is a measure of how uncertain the model is about its response. It is measured as the exponent of the response's cross-entropy  (i.e. the negative mean of the output tokens' conditional log probabilities.\footnote{See \citet{jm2}.}) A perplexity of 1 indicates that the model has no uncertainty about its output, while a higher perplexity means the model is more uncertain.  % NO LONGER USING PERPLEXITY

%\paragraph{Agenda order.} We also hypothesize that the order in which a case appears in the agenda may matter for its outcome. We use the order in which the case appears in the agenda, not the order in which the case was discussed at the actual meeting. The committee chair has the ability to discuss agenda items out of order, and this may be endogenous to the meeting outcomes, so we focus on the order in which the item appears in the agenda published prior to the start of the meeting.

%\paragraph{Number of agenda items.} We hypothesize that the number of items on the agenda can have an effect on outcomes; in particular, cases are more likely to be postponed if there are a large number of items on the agenda.

%\paragraph{Consent calendar.} The Los Angeles CPC utilizes a practice for streamlining meetings known as the ``consent calendar''. The consent calendar takes multiple agenda items and groups them into a single motion that the committee votes on together as a whole. The consent calendar tends to include cases that the committee chair has deemed non-controversial and therefore not requiring separate discussion. The consent calendar is published \emph{before} the meeting starts, and items can be taken off the consent calendar during the meeting. We hypothesize that being on the consent calendar significantly predicts approval.

%\paragraph{Number of support and opposition letters.} We hypothesize that the amount of public support and public opposition matters for hearing outcomes. Greater public support is expected to improve a proposal's probability of being approved, whereas greater public opposition increases the likelihood that the proposal is denied, delayed, or approved with conditions. We include public support and public opposition as two separate explanatory variables to allow for heterogeneous impacts of support vs. opposition. Note that a proposed project can receive multiple letters both in support and in opposition. We measure public support as the log base 2 of the number of letters written in support, and we measure public opposition as the log base 2 of the number of letters written in opposition, as discussed in Section \ref{sec_data}. Figure \ref{fig_support_oppose} shows the distribution of the number of letters in support and in opposition across cases. 

%\paragraph{Council districts.} Council districts may matter for case outcomes for a variety of reasons. For one, although City Planning Commission members are appointed from a variety of professional backgrounds, most of them come from backgrounds of urban planning, public service, real estate development, or community advocacy. A CPC member, despite best efforts to remain impartial, may therefore still be influenced by the specific politics of the council district. For another, the Los Angeles City Council has veto power over City Planning Commission decisions, and the Council usually defers to the member whose district the project is located in for such decisions. CPC members may therefore be cognizant of the opinions of the project district's council member when making their decisions. To control for these possibilities, we include council district fixed effects as explanatory variables.

%\paragraph{Case suffixes.} As discussed in Section \ref{sec_data}, case suffixes indicate the types of entitlements requested or required by the project, as well as other project indicators. In the data, there are \gn{NumUniqueSuffixes} unique suffixes, some of which are quite sparsely populated, which is problematic for our dataset of \gn{NumberOfCases} observations. We therefore collapsed the suffixes into \gn{NumSfxGroups} logical groups with fixed effects for each. The largest suffixes by frequency were retained as their own buckets, and the remaining were grouped according to similarity in their legal descriptions and procedural functions. The suffixes were grouped into the following categories:
%\input{results/suffix_groups.tex}









