\section{Results}\label{sec_results}

Table \ref{tab_ologit_results} reports the results from the ordered logit regression. The main coefficient of interest is the effect of case atypicality on case outcome. Four specifications are presented which show the effect of progressively adding more explanatory variables to the regression. The coefficient on case atypicality is robust to the gradual inclusion of more and more confounding variables. Importantly, by including suffix, embedding cluster, and council district fixed effects, we show that the effect of atypicality on case outcome is not driven by selection on case type, category of requested entitlement, or geography.  

%We estimate an ordered logit model to analyze the determinants of project outcomes. The dependent variable is increasing in approval status as proposals can be rejected or delayed, approved with conditions or modifications, or granted full approval. All models are estimated with robust standard errors. We progressively add fixed effects for embedding cluster, council district, and suffix groups to account for heterogeneity in proposal types, jurisdiction, and administrative classifications. Table \ref{tab_ologit_results} reports the results. 

%The results suggest that evaluating proposals is cognitively demanding, and unusual ones are harder to process. 
%We operationalize a proposal's recognizability through case atypicality, a measure of how closely it resembles the kinds of cases commissioners typically encounter. Higher values indicate a less typical proposal. 
The coefficient on case atypicality is negative and statistically significant across specifications, indicating that unusual cases are less likely to be approved outright, and more likely to be approved with conditions or delayed or denied. Based on our preferred specification in column 4 of Table \ref{tab_ologit_results}, a one-standard-deviation increase in atypicality decreases the log odds of full approval by \gn{SemUniCoef} ($p<0.05$). By contrast, being placed on the consent calendar increases the log odds of full approval by \gn{ConCalCoef} ($p<0.01$). These results demonstrate the measurable penalty that unusual projects face from imposing higher cognitive costs on commissioners. Meanwhile, routine consent calendar items move through the process almost automatically. Other institutional factors such as agenda order and agenda length are small and statistically insignificant, indicating that bureaucratic frictions may stem less from meeting logistics than from project familiarity or project typicality.

Commissioners also weigh the effects of opposition on reputational and political exposure. The results indicate that doubling the number of opposition letters decreases the log odds of approval category by \gn{NOppCoef} ($p<0.01$). In contrast, the coefficient on support letters is small and statistically insignificant. Commissioners appear systematically more sensitive to community opposition than to community support, perhaps reflecting a defensive orientation in their decision-making.

Table~\ref{tab_ologit_marginal_effects} reports the marginal effects, which can be interpreted as the predicted effect on outcome probability for a one unit change in the regressor, averaged over cases. 
%Moreover, the patterns are monotonic: outcomes worsen as atypicality and opposition increase, and improve across outcomes for consent calendar items. 
The results indicate that a one standard deviation increase in atypicality reduces the average probability of full approval by \gn{SemUniME2}\% ($p<0.05$), while raising partial approval by \gn{SemUniME1}\% ($p<0.05$) and denial/postponement by \gn{SemUniME0}\% ($p<0.05$). Likewise, for opposition letters, doubling their number lowers the average likelihood of approval by \gn{NOppME2}\% ($p<0.01$), and increases partial approval by \gn{NOppME1}\% ($p<0.01$) and denial/postponement by \gn{NOppME0}\% ($p<0.01$). Placement on the consent calendar has the opposite effect. Full approval increases by \gn{ConCalME2}\% ($p<0.01$), partial approval decreases by \gn{ConCalME1}\% ($p<0.01$), and denial/postponement decreases by \gn{ConCalME0}\% ($p<0.01$).

To assess the robustness of our findings to omitted variable bias, we follow the approach of benchmarking selection on unobservables against selection on observables \citep{AltonjiElderTaber2005, Oster2019}. If adding a rich set of controls barely shifts the coefficient of interest while substantially increasing explanatory power, then unobservables would have to be implausibly more important than observables to overturn the result. Using the \citet{Oster2019} framework with $R_{\text{max}} = 1.3 \tilde{R}$, we calculate $\delta^\ast = \gn{OsterDelta}$, which well exceeds the recommended threshold of $\delta=1$. This means that to reduce the coefficient on case atypicality to zero would require selection on unobservables to be about three times as as important as the included controls.

%We use movements in coefficients and $R^2$ across nested specifications to compute the relative strength unobservables would need to eliminate the effect \citep{Oster2019}. Given standard benchmarks, our estimates are highly robust to concerns about omitted variable bias.

Overall, these results are consistent with our hypothesis that bureaucratic decision-making at the CPC reflects both bounded rationality and oversight risk. Decisions on unfamiliar projects may be postponed because they require greater cognitive processing, while contested projects are also put off because they expose commissioners to greater political and reputational scrutiny. By contrast, routine and uncontested projects proceed with little resistance. These patterns highlight how the regulatory process itself shapes outcomes as a function of the trade-offs faced by commissioners.

\subsection{Discussion: what happens to delayed cases?}

We have argued that atypicality and public opposition are both salient factors that increase the risk of a proposal getting delayed at the CPC. Two important questions remain: 
\begin{enumerate}
\item Are delayed cases more likely to eventually be denied? 
\item And, what is the impact of delay on overall development?
\end{enumerate}
The first question pertains to whether the effect of atypicality is due to cognitive frictions (as we argue), or due to systematically lower quality of atypical proposals. If atypical proposals are also of systematically lower quality, then we would expect them to be significantly more likely to be denied in subsequent hearings, or only approved with conditions. To test this hypothesis, we remove all delay decisions from the estimation sample and estimate a binary logit model in which the outcomes are 1 if the case is approved and 0 if the case is either approved with conditions or denied. The results of this regression are shown in Table \ref{tab_blogit_results}. Once all the other regressors are included (column 4), the estimated influence of case atypicality becomes statistically insignificant. The coefficient is still positive, suggesting that atypical cases may still be more likely to be ultimately denied or only approved with conditions, but the effect is weaker in both statistical significance and economic magnitude in comparison to the effect of atypicality on delay.

The second question pertains to whether delays at the CPC have a meaningful impact on economic outcomes. \citet{gabrielkung2025} argued that approval delays are an important driver of reduced housing production. They estimated that a 25\% reduction in approval times would have increased housing production in Los Angeles by 12.7\% between the years of 2010 and 2022. Thus, delays at the CPC could have a meaningful impact on real estate development if they add significant time to the overall approvals process. Table \ref{tab_delayed_outcomes} investigates what happens to the delayed cases in our dataset. There were \gn{NumDelayedCases} cases which were ever delayed in our dataset.\footnote{This number differs from the \gn{CasesContinued} instances of delay reported in Table \ref{tab_result_unanimity} because Table \ref{tab_result_unanimity} counts hearing outcomes and some cases can be heard multiple times within our dataset.} Of these, \gn{NumResolved} were ultimately voted on within our dataset. The average time between first observed delay and final outcome within the data was \gn{DaysToResolve} days. This is an underestimate because a) we don't observe the final outcome for some of the cases (right-censored resolution times) and b) we don't necessarily observe the first hearing (left-truncated start times). But even this underestimate of \gn{DaysToResolve} days is economically significant: using data from Los Angeles from 2010 to 2022, \citet{gabrielkung2025} report that the median approval time for multifamily housing development projects was 524 days, so \gn{DaysToResolve} days is a meaningful addition. Moreover, it is interesting to note that between the initial delay and final outcome, the majority of the delayed cases (\gn{ResolvedNoChangesPct}\%) did not change in terms of their requests, and some (\gn{ResolvedMinorChangesPct}\%) came back with only minor changes. Only \gn{ResolvedMajorChanges} of the delayed cases came back with any major changes to their requests; and even amongst these six, only \gn{ResolvedMajorReduce} came back with reduced requests, while the other \gn{ResolvedMajorExpand} actually expanded their requests. This suggests that the development impact of CPC delay operates primarily through a time-delay channel, rather 

