\section{Conclusion}\label{sec_conclusion}

%This paper examined how cognitive and oversight constraints shape bureaucratic decision-making in the Los Angeles City Planning Commission. We developed a text-based measure of case atypicality to capture the cognitive effort that unusual proposals impose on commissioners, and used sentiment analysis of public correspondence to measure monitoring pressure. The results reveal that both constraints have significant effects on regulatory outcomes. Less familiar projects, which require greater interpretive effort, are less likely to receive full approval, while routine items proceed with little resistance. Public opposition exerts a similarly negative effect, whereas public support provides little offsetting influence. These patterns provide empirical evidence of bounded rationality in administrative decision-making, with bureaucrats allocating scarce attentional resources across proposals, prioritizing those that are familiar or safe under scrutiny. Institutional mechanisms such as the consent calendar function help economize on attention.

This paper brought structure to the public record of the Los Angeles City Planning Commission, converting eight years of agendas, minutes, and public correspondence into a systematic dataset of 818 development decisions. That record reveals clear patterns. The Commission's operative margin is delay, and delayed cases typically return unchanged after an average of 105 days. Approval odds fall with case atypicality---a text-based measure we developed to capture how unusual a proposal is relative to the docket the Commission typically hears---and with public opposition, whereas public support provides little offsetting influence. Routine items, particularly those bundled onto the consent calendar, proceed with little resistance. These patterns are consistent with commissioners allocating scarce attentional resources across proposals under public scrutiny, prioritizing cases that are familiar or safe. Institutional mechanisms such as the consent calendar function help economize on attention.

The results also have implications for urban development. Land-use regulation remains one of the central bottlenecks in housing supply, yet little is known about the role of the bureaucratic process vis-\`a-vis the statutory requirements. Our analysis shows that administrative capacity itself may constrain housing production beyond the direct impact of statutory restrictions. The magnitudes are meaningful: the average 105-day delay we observe at the CPC stage is a substantial addition to the 524-day median approval timeline documented for multifamily projects in Los Angeles (\citealp{gabrielkung2025}). Negative public pressure clearly shapes outcomes, but cognitive limits---and thus, bureaucratic capacity---also constrain approval of novel or complex projects. The result is a form of institutional lock-in. By favoring projects that are easiest to interpret and defend, the process may reinforce its own expectations, narrowing the range of designs it can evaluate, and thereby limiting innovation in the kinds of development that could meet residents' needs.
%The outcome may be an urban landscape optimized for administrative legibility rather than public value.

More broadly, our framework demonstrates how text can be used to represent the decision architecture within organizations. With such data becoming increasingly available, atypicality provides a scalable measure to study how information is processed, attention is allocated, and discretion is transformed into routines. These mechanisms deepen our understanding of how organizations adapt, and the conditions under which they resist change.

\pagebreak