AYES:\\ 
<A comma separated list of commission members who voted for the motion. If multiple motions were made, use only the last motion. If no one voted for, say "NONE". If no motion was made, say "N/A".>\\ 
\\ 
NAYS:\\ 
<A comma separated list of commission members who voted against the motion. If multiple motions were made, use only the last motion. If no one voted against, say "NONE". If no motion was made, say "N/A".>\\ 
\\ 
ABSTAINED:\\ 
<A comma separated list of commission members who abstained from voting on the motion. If multiple motions were made, use only the last motion. If no one abstained, say "NONE". If no motion was made, say "N/A".>\\ 
\\ 
RECUSED:\\ 
<A comma separated list of commission members who were recused from voting on the motion. If multiple motions were made, use only the last motion. If no one was recused, say "NONE". If no motion was made, say "N/A".>\\ 
\\ 
ABSENT:\\ 
<A comma separated list of commission members who were absent. If multiple motions were made, use only the last motion. If no one was absent, say "NONE". If no motion was made, say "N/A".>\\ 
\\ 
VOTE RESULT:\\ 
<Did the motion pass or fail? Your only options are MOTION PASSED, MOTION FAILED, N/A>\\ 
\\ 
RESULT OF APPEAL:\\ 
<If the agenda item involved an appeal, say whether the appeal was granted or denied. Your only options are: APPEAL GRANTED, APPEAL GRANTED IN PART, APPEAL DENIED, NO APPEAL, DELIBERATIONS CONTINUED TO FUTURE DATE, APPLICATION WITHDRAWN>\\ 
\\ 
IMPLICATION FOR PROJECT:\\ 
<What is the implication of the vote for the original requested actions? Your only options are: APPROVED, APPROVED IN PART OR WITH MODIFICATIONS, DENIED, DELIBERATIONS CONTINUED TO FUTURE DATE, APPLICATION WITHDRAWN>